---
title: "Visualizing with D3 JavaScript Network Graphs from R"
author: "John Letteboer"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    highlight: haddock
    keep_md: yes
    theme: cosmo
---

Hi all, I want to visualize some of my courses I've followed by year. I have found [D3 JavaScript Network Graphs from R](http://christophergandrud.github.io/networkD3/) with some nice and cool visualizations.

In this post I will describe how I transform my data for creating the RadialNetwork and DiagonalNetwork visualizations.

# Loading and have a look at the data
First off all I had a Excel file with some of my followed some courses since 2013. I've saved it to a csv file ```courses.csv```. Let have a look of the file.

```{r}
setwd("~/Documents")
courses <- read.csv("courses.csv", sep=";",header=TRUE, stringsAsFactors = FALSE)
head(courses)
str(courses)
```

Oke, we have a data.frame of 25 observation of 4 variables (year, month, course and type). Since both visualizations are based on tree diagrams we need to create a tree from my ```data.frame```. After some searching on the internet I've found the package [data.tree](https://cran.r-project.org/web/packages/data.tree/vignettes/data.tree.html). In this package there is a paragraph called "Create a tree from a ```data.frame```".

# Let's start to create a tree from my courses
Load the library.
```{r}
library(data.tree)
```

The ```data.frame``` is a table and each row is a leaf. We create a path from root to leaf and save that into the ```data.frame``` as new column ```pathString```.
Define the hierarchy (Year|Month|Type|Course)
```{r}
courses$pathString <- paste("courses",courses$year,courses$month,courses$type,courses$course, sep= "|")
head(courses)
```

After we have created the ```pathString``` we need to convert it to a node, with the package ```data.tree``` is that very easy.
```{r encoding = 'UTF-8'}
coursesTree <- as.Node(courses, pathDelimiter = "|")
coursesTree
```

The last thing we need is a list-to-list structure before we can plot the two networkD3 visulizatios
```{r}
coursesTreeList <- ToListExplicit(coursesTree, unname =TRUE)
```

# Plot the two visulizatios
```{r fig.width=9, fig.height=8}
library(networkD3)
diagonalNetwork(List=coursesTreeList, 
                fontSize = 13, 
                fontFamily = "OpenSans-Light", 
                nodeStroke = "orange",
                linkColour = "#AAA",
                opacity = 0.9)

radialNetwork(List=coursesTreeList, 
                fontSize = 13,
                fontFamily = "OpenSans-Light",
                nodeStroke = "orange",
                opacity = 0.9)
```

Well that look nice!! This package is great!!

This code can also be find on my [GitHub](https://github.com/jletteboer/networkx.nl/)

## SessionInfo
```{r}
sessionInfo()
```
