---
title: "Create blog posts from RStudio to Wordpress"
author: "John Letteboer"
date: "12/14/2015"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    theme: cosmo
---

In a couple of courses I've used [RPubs](http://www.rpubs.com/) and [GitHub](https://github.com/) for publishing my [R](https://www.r-project.org/) code and work. As I have a own website with [Wordpress](https://wordpress.org/) I want to blog from R to my own blog instead only through sharing R code via RPubs and GitHub. After searching over the internet I find some usefull stuff to do this. So let's try it myself en let this be my first blog post created from [Rstudio](https://www.rstudio.com/) to Wordpress.

### Setting up Wordpress and R
##### What do we need for wordpress:  
- [Crayon Syntax Highlighter](https://nl.wordpress.org/plugins/crayon-syntax-highlighter/) plugin for highlighting R code in a blog post.  
- [WP QuickLaTeX](https://nl.wordpress.org/plugins/wp-quicklatex/) for nice math rendering using $L^AT_EX$ syntax.  

##### And in R we need:  
- [knitr](http://yihui.name/knitr/), a website is coded in HTML, `knitr` can convert your R code from a [R Markdown](http://rmarkdown.rstudio.com/) file to HTML code.this package is already installed with RStudio.  
- [RWordPress](http://yihui.name/knitr/demo/wordpress/) allows publish blog posts from R to WordPress.  

##### Load the necessary packages in R
```{r}
if (!require('RWordPress'))
  install.packages('RWordPress', repos = 'http://www.omegahat.org/R', type = 'source')
library(RWordPress)
library(knitr)
```

Setting options for RWordPess
```{r}
options(WordpressLogin=c(your_username='your_password'),
        WordpressURL='http://your.blog.com/xmlrpc.php')
```

##### Syntax highlighting with Crayon Syntax Highlighter
To make syntax highlighting work in WordPress with the Crayon Syntax Highlighter plugin. R code should be enclosed in WordPress-shortcode instead of the knitr html output default 
```html
<pre><code class="r">...</code></pre>
```
to 
```html
<pre class="lang:r decode:true">...</pre>
```

We can't do that with the default knitr2wp function of knitr. `knitr2wp` is only based on the syntax which is used for [SyntaxHighlighter](https://wordpress.org/plugins/syntaxhighlighter/). I don't want to update all my other code on my blog.

For the Crayon Syntax Highlighter I've found a [request](http://git.leg.ufpr.br/leg/legTools/commit/4e76bb02f76cbf9f11b3ddaedd73421536bc94ac) of allowing proper code highlight when using Crayon. Below the copied requested function for using it to post a blog  to my Wordpress site with Crayon Syntax Highlighter as code highlighter.
```{r}
knit2wpCrayon <- function(input, title="A post from knitr", ...,
                          action=c("newPost", "editPost", "newPage"),
                          postid, encoding=getOption("encoding"),
                          upload=FALSE, publish=FALSE, write=TRUE)
{
    out <- knit(input, encoding=encoding)
    on.exit(unlink(out))
    con <- file(out, encoding=encoding)
    on.exit(close(con), add=TRUE)
    content <- knitr:::native_encode(readLines(con, warn=FALSE))
    content <- paste(content, collapse="\n")
    content <- markdown::markdownToHTML(text=content, fragment.only=TRUE)
    content <- gsub("<pre><code class=\"([[:alpha:]]+)\">(.+?)</code></pre>",
                    "<pre class=\"lang:\\1 decode:true\">\\2</pre>",
                    content)
    content=knitr:::native_encode(content, "UTF-8")
    title=knitr:::native_encode(title, "UTF-8")
    if (write){
        writeLines(text=content,
                   con=gsub(x=out, pattern="\\.md$", replacement=".html"))
    }
    if (upload){
        action=match.arg(action)
        WPargs=list(content=list(description=content, title=title, 
                                 ...), publish=publish)
        if (action=="editPost") 
            WPargs=c(postid=postid, WPargs)
        do.call("library", list(package="RWordPress", character.only=TRUE))
        print(do.call(action, args=WPargs))
    }
}
```

After adding the function into R we can post our first blog on Wordpress and that would be this one :). 

### Publishing the post to your blog

**Note:** Below code should run directly from the R shell (first create the above function and load the RWordPress library) after the R Markdown has created and has runs once. It will post a lots same drafts and ends with a error. I cannot figure out why, I think there is a loop and it wants to post itself to your blog. If you want to upload it run it directly from the R shell with the upload option to `TRUE`.

```{r}
#knit2wpCrayon("r2blog.Rmd", 
#        title = "Create blog posts from RStudio to Wordpress",
#        categories = c("R", "Programming"), 
#        publish = FALSE, upload = FALSE)
```

I set the publish option to `FALSE`, so the post will stay in draft mode and I can fine tune some formats, tags, set a featured image and add on the top the shortcode `[latexpage]` for using $L^AT_EX$ syntax. 

This code can also be find on my [GitHub](https://github.com/jletteboer/networkx.nl/tree/master/r2blog)